\documentclass[a4paper,12pt]{report}

\usepackage{listings}
\usepackage{xcolor}
\usepackage{geometry}
\geometry{margin=1in}

\lstset{
    basicstyle=\ttfamily\small,
    keywordstyle=\color{blue}\bfseries,
    commentstyle=\color{green!50!black},
    stringstyle=\color{orange},
    showstringspaces=false,
    frame=single,
    breaklines=true
}

\begin{document}

\begin{center}
    \Huge \textbf{DBMS Practical File} \\
    \vspace{1cm}
    \Large Practical – Implementation of Strict 2PL and Rigorous 2PL \\
    \vspace{3cm}
    \textbf{Name: Anshuman Diwakar} \\
    \textbf{Course: B.Tech (CSE)} \\
    \textbf{Subject: Database Management Systems} \\
    \textbf{Instructor: Karan Gupta} \\
    \textbf{Date: September 9, 2025} \\
\end{center}

\newpage

\section*{Aim}
To demonstrate \textbf{Strict 2PL} and \textbf{Rigorous 2PL} protocols using SQL transactions.

\section*{Theory}
\begin{itemize}
    \item \textbf{Strict 2PL}:
    \begin{itemize}
        \item Shared (S) locks may be released after use.
        \item Exclusive (X) locks are held until commit/abort.
        \item Ensures conflict-serializable and cascadeless schedules.
    \end{itemize}

    \item \textbf{Rigorous 2PL}:
    \begin{itemize}
        \item Both shared and exclusive locks are held until commit.
        \item Ensures conflict-serializable, cascadeless, and strict schedules.
    \end{itemize}
\end{itemize}

\section*{Database Setup}
\begin{lstlisting}[language=SQL]
CREATE TABLE Accounts (
    acc_no INT PRIMARY KEY,
    balance INT
);

INSERT INTO Accounts VALUES (1, 1000);
INSERT INTO Accounts VALUES (2, 2000);
\end{lstlisting}

\section*{Transaction 1 (T1)}
\begin{lstlisting}[language=SQL]
BEGIN TRANSACTION;

-- T1 reads Account 1
SELECT balance FROM Accounts WHERE acc_no = 1;

-- T1 updates Account 2
UPDATE Accounts SET balance = balance + 500 WHERE acc_no = 2;

-- Commit releases locks
COMMIT;
\end{lstlisting}

\section*{Transaction 2 (T2)}
\begin{lstlisting}[language=SQL]
BEGIN TRANSACTION;

-- T2 updates Account 1
UPDATE Accounts SET balance = balance - 200 WHERE acc_no = 1;

-- T2 reads Account 2
SELECT balance FROM Accounts WHERE acc_no = 2;

COMMIT;
\end{lstlisting}

\section*{Schedule under Strict 2PL}
\begin{verbatim}
T1: S-lock(A1), read(A1), release S-lock(A1)
T2: X-lock(A1), write(A1) (waits until T1 releases)
T1: X-lock(A2), write(A2), commit (releases X-lock)
T2: read(A2), commit
\end{verbatim}

\section*{Schedule under Rigorous 2PL}
\begin{verbatim}
T1: S-lock(A1), read(A1) (lock not released)
T2: waits for X-lock(A1)
T1: X-lock(A2), write(A2)
T1: commit (all locks released)
T2: X-lock(A1), write(A1), S-lock(A2), read(A2), commit
\end{verbatim}

\section*{Conclusion}
\begin{itemize}
    \item \textbf{Strict 2PL} holds exclusive locks till commit but may release shared locks earlier.
    \item \textbf{Rigorous 2PL} holds \emph{all locks} (S and X) till commit.
    \item Rigorous 2PL gives the highest degree of isolation and avoids cascading aborts completely.
\end{itemize}

\end{document}
